---
title: "ISTA 320 Data Visualization Challenge 7"
author: "Tiamara Craig"
date: "Spring 2022"
output: html_document
---


```{r setup, include=FALSE}
library(knitr)
library(tidyverse)
library(dbplyr)
opts_chunk$set(echo = TRUE)
```

```{r}
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

# Read data in
```{r}
work_data <- read_csv("data/work.csv", show_col_types = FALSE)
```

# filter by country ie United States
```{r}
USA_Data <- filter(work_data,Country =='United States')
```

# drop unneeded / redundant columns / update col name "Value" to "Hours Weekly" and "Earnings" to "Wage"
```{r}
colnames(USA_Data)[13] <- "Hours_Weekly"
colnames(USA_Data)[5] <- "Wage"
clean_USA_Data <- USA_Data[ c(3,5,12, 13) ]

```

# group and sort the col's by year and store it in a new data frame
```{r}
new_clean_USA_Data <- clean_USA_Data %>%
  group_by(Year) %>%
  arrange(desc(Year)) %>%
  group_by()
```


# Get rid of duplicate rows and filter out wages that are not Minimun wage.
```{r}
uniq_USA_data <-new_clean_USA_Data[!duplicated(new_clean_USA_Data), ]
final_USA_Data <- filter(uniq_USA_data, Wage =="MIN")
```

# Plot #1: Bar plot (colorblind frienldly) acc to http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/
```{r}
bp <- ggplot(data = final_USA_Data, mapping = aes(x=Year, y=Hours_Weekly, fill=FAMILY)) + geom_col(stat="identity", position = "dodge")
bp +
  scale_fill_manual(values=cbPalette) + ggtitle("Weekly Hours to Work to Stay Sbove Poverty Line") +  scale_x_continuous(breaks = scales::pretty_breaks(n = 10))

```

# Plot #2: Scatter plot 
# create a new data frame that shows the avg hours per year
```{r}
Yearly_Avg_Data <- final_USA_Data[ c(3,4) ]

grouped_Year_Data <- Yearly_Avg_Data %>%
  group_by(Year) %>% 
  summarize(avg=mean(Hours_Weekly))

sp <- ggplot(data = grouped_Year_Data, mapping = aes(x = Year, y = avg)) +
  geom_point()  + ggtitle("The Average Hours Needed to Stay Above the Povety Line Each Year")+  scale_x_continuous(breaks = scales::pretty_breaks(n = 20)) + geom_smooth()  +ylab("Avg Hours per week")

sp

```


# Plot #3: Scatter plot seperated
# using final_USA_Data data set
```{r}
gp <- ggplot(data = final_USA_Data, aes(x = Year, y = Hours_Weekly)) +
    geom_line() +
    facet_wrap(facets = vars(FAMILY)) 
 gp  + ggtitle("The Average Hours Needed to Stay Above the Povety Line Each Year for each Family Type") +
  scale_fill_manual(values=cbPalette)
 
```  
